---
title: >
  Reproduction of Eglen and Wong (2008) --- Spatial constraints
  underlying the retinal mosaics of two types of horizontal cells in
  cat and macaque
author: Stephen J. Eglen (ORCID 0000-0001-8607-8025)
output:
  pdf_document:
    toc: false
    includes:
       in_header: preamble.sty
bibliography: refs.bib
---

```{r rsetup,eval=TRUE,include=FALSE}
require(knitr)
options(width=60)
opts_chunk$set(cache=FALSE)
```

## Summary

This paper reports the successful reproduction of the results in a
paper I co-authored in 2008 [@Eglen2008-ca].  A postprint of this
paper is available at <https://arxiv.org/abs/1208.0986>.  (Open access
was not mandated by the funders of this research at the time of
publication, 2008.)  All the code has been uploaded to
<https://github.com/sje30/rescience-hor> along with the Rmarkdown
sources for this report.


<!-- <https://github.com/ReScience/ten-years/blob/master/author-guidelines.md> -->

# Historical context

## Scientific question

To investigate a problem in neurobiology concerning the relative
spatial positioning of two types of neuron in the retina.  We used
three example datasets (pretty much all that was available at the
time) to investigate this problem using spatial statistics and
simulations of spatial patterning.  These datasets were not available
in digital format, but the images [@Wassle1978-cd; @Wassle2000-aa]
were scanned in and digitised (each cell represented as an x,y
location) by me as part of the work for the original paper.  These
datasets were not publicly shared before.


## Computational context

Most of the code required for the 2008 paper had been developed by
myself since about 2000 working on other related problems.  The major
innovation here was in the biological problem.

One of the key reasons I switched to R as a programming environment in
the late 1990s was because of the strong support it had for spatial
statistics.  In particular the SPLANCS package [@splancs] in R has
been a key package in most of my research.

## Original source code

the original source code was not published or publicly archived
before.  The source code was kept just on my hard drive.  I keep good
backups though, and in this case the projects were still in my home
directory (copied from machine to machine over the years).  No license
was written systematically to cover the code I wrote, although the R
packages are usually released under GPL or MIT.

As a side-effect of producing this article, the following R packages
of mine were made available:

1. sjedmin:  dmin model for spatial point patterns.
   <https://github.com/sje30/sjedmin>
2. sjedist: unified calculation of spatial statistics.
   <https://github.com/sje30/sjedist>
3. sjevor: Wrapper around public domain code for Voronoi tessellation
   <https://github.com/sje30/sjevor>
   
The Voronoi tessellation was public domain C code provided by Steve
Fortune [@Fortune1987-kg].  (This code was not required for the current
project, but was implicitly required for the parent package, sjedist.)

## Retrieval of the software

This was relatively straightforward, as the software was all on my
hard drive.  One complication was that until a few years ago, I had
been using RCS as my primary version control software.  As I wished to
release the packages onto github, I converted the history from RCS into
Git using the tool `https://github.com/Oblomov/rcs-fast-export`.


I anticipated that loading the data might be tricky, as I had used
R's internal Rda format to efficiently store the data files (they are
not large, so there was no need to be concerned about efficiency).
Given that this format could in principle change over time, this might
have been a problem.  However, R 4.0.0 was able to read in the file
`bivariate_mosaics.Rda`.

# Execution

The code ran fine in today's version of R (4.0.0); see end of document
for the full list of R dependencies.  I did not need to load any
older versions of libraries.  The code was tested on both my laptop
(ThinkPad T480s running Manjaro linux) and on <https://mybinder.org> (see the
end of this document for mybinder details.)   The simulations
themselves only take a couple of minutes to run.  If recompiling on
binder, it takes a few more minutes to install all the packages and
initialise the tex distribution.

However, although the code worked fine, it took me about 20 minutes to
read through the old scripts and work out what to do, as I had not
provided a Makefile or similar.

Once the code was running, I only wished to make cosmetic updates to
the project to modernise it.  The most significant change was
converting the  format for figures from postscript to pdf as there is
more support for the latter.  The three figures that were generated
are reproduced in this paper and can be compared against the figures
from the original paper.

A Makefile is provided to help regenerate the simulation results and
the figures.  *make clean* will return the repository to a clean
state.  *make all* will first run the simulations and store the
results in several *.Rda* files.  There are then several R scripts to
generate the figures from the corresponding *.Rda* files.  Finally,
after the figure pdfs have been generated, this document is compiled
from its Rmarkdown source.

## Reproduction of results

The results generated closely matched those that were presented in the
original paper.  As the simulation methods are stochastic (and no seed
was set for reproducible random number generation), the P values
reported in Figure 2 do not match those originally reported, nor does
the simulation field (right hand side of Figure 1) match.  However,
the statistical properties of the models still match those of the
experimental data.

## Could someone else reproduce the work?

If I had simply given the R code and packages to someone else, the
lack of documentation or Makefile would have meant that this person
would have needed to study the code carefully to work out what files
were generated.  I would have hoped though that someone fluent with R
would have made sense of the code itself, and all of the parameters
for the simulations were included in the source files.


# Reproduction of Figure 1

\includegraphics{hor_fieldA_col.pdf}

We also made a supplementary figure of all 3 fields.

# Reproduction of Figure 2

\includegraphics{eglen_fig2_col.pdf}

# Reproduction of Figure 3

\includegraphics{eglen_fig3_col.pdf}


# Reproduction of Table 2

Key column from table 2 is the rejects column which is shown here:

```{bash, code = readLines("table2-rejects.txt"), eval=FALSE,echo=TRUE}
```

Note the order of the three values is field A, B, C, rather than
listed in the original paper (Table 2, C, A, B).  The other columns of
this table were determined statically from the data sets, so were not
recalculated here -- the rejects column is the important column here.

(Had I known more about generating tables in R at the time, I think
my R code should have created this entire table.)

# About this document

This is a reproducible document that you can regenerate within
mybinder.org by following the instructions on the README at the
project home page, <http://github.com/sje30/rescience-hor>.

If you are not using mybinder.org to recompile the document, then the
following steps should work to generate the file `rescience-hor.pdf`
on a unix machine with Make.

```{r eval=FALSE}
source('install.R')
system('make rescience-hor.pdf')
```


\clearpage
# Session Information

```{r}
sessionInfo()
```

# Acknowledgements

I thank the R core team for continuing to provide a robust computing
environment that gave me confidence that this code was still work
after 10 years.  Thanks also to Daniel Nüst and the mybinder team for
helping me learn and use mybinder.


\clearpage
# References

